Stack Safety is a Security Property

### 2021-04-10 tex/main.tex

It may be unclear in figure \ref{fig:conf1} why we also vary the initial values of {\tt w}
and {\tt z}. We will demonstrate by amending our toy model to make it closer to a real stack.
In figure \ref{fig:conf2} the topmost stack frame is no longer allocated as needed, but is an
infinite array of addresses. Variables are allocated and deallocated by moving the stack pointer,
and on a call, addresses to the left of the stack pointer are split off and pushed onto the
stack as sealed frames. On a return, the topmost sealed frame is prepended to the unsealed
frame. Now we can see how secrets might be left behind even in the unsealed portion of the stack.

\sna{Updated up to here.}

Two important details of confidentiality are reflected in figure \ref{fig:conf2}.
First of all, it has multiple calls, and while we focus on the stack at the call
from line 13 to line 20, it should be clear that we also need to consider a similar
variant at the call from 0 to 10. Confidentiality is in this sense recursive: each
nested call must obey confidentiality in its own right.

More importantly, careful readers will have noticed that the informal definition
of stack confidentiality refers to a variation of the entire stack, not just the
sealed portion. In fact, we support temporal stack safety, in which values left behind
by previous activations are also inaccessible. In this case, we connect temporal safety
firmly to the caller-callee attacker model -- {\tt f} stored some data in the variables
{\tt y} and {\tt z}, used it, then deallocated them by moving the stack pointer.\apt{Where is the storage for x ?.
And need to explain what the colors mean.}
After the stack pointer is written over {\tt y}, the value of {\tt z} is still
left behind. Then {\tt g} can read it just by allocating and accessing an uninitialized
variable! \apt{Need to be more explicit about what is getting varied, when, and why.}

\paragraph*{Control Flow and Awkward Examples}

\begin{figure}
  \centering
  \begin{minipage}{.4\textwidth}
\begin{verbatim}
int stash = 0;

int main() {
0:     int x = 1;
1[C]:  f();
2:     x = -x;
3[C]:  f();
4[R]:  return x;
}
\end{verbatim}
  \end{minipage}
  \begin{minipage}{.4\textwidth}
\begin{verbatim}
void f() {
10:     if(!stash) {
        // store return pointer
11:       stash = $SP-1;
12:     } else {
13:       $SP-1 = stash;
14:       stash = 0;
15[R]:    return;
        }
}
\end{verbatim}
  \end{minipage}
\label{fig:WBCF1}
%  \begin{subfigure}[t]{.4\textwidth}
%\begin{verbatim}
%int counter = 0;
%int stash = 0;
%
%int main() {
%0:     int x = 1;
%1[C]:  f();
%2:     x = -x;
%3[C]:  f();
%4[R]:  return x;
%}
%
%void f() {
%10:     switch(counter++) {
%11:       case 0:
%12[R]:      return;
%13:       case 1:
%14:         stash = $SP-1;
%15[C]:      main();
%16:         return;
%17:       default:
%18:         $SP-1 = stash;
%19[R]:      return;
%        }
%}
%\end{verbatim}
%\label{fig:WBCF2}
%  \end{subfigure}
  \caption{Control Flow Attacks}
  \label{fig:WBCF}
\end{figure}

Example \ref{fig:WBCF} shows how control flow attacks can compromise integrity in principle
even when the attacker obeys LSE. In this case, the attacker hides a return pointer in a
global variable and retrieves it the second time it is called, returning to the first call
point rather than the second. A return was performed, so {\tt x} is no longer sealed, but
line {\tt 2} is executed twice, so that {\tt main} returns 1 instead of -1.

To prevent this, we define Well-bracketed Control Flow. Throughout the above examples, we
have implicitly tracked the depth of the current call in order to seal memory at the correct
depth. We also have a clear sense of where the entry points to functions are -- the top line,
in our examples, and another instruction annotation in practice. Well-bracketed Control Flow
holds if:

\begin{itemize}
\item All calls jump to entry points ({\em entry integrity})
\item The depth counter decreases by exactly one on a return ({\em non-skipping})
\item From a call at depth \(d\), the next state to have depth \(d\) matches the stack pointer,
  and the program counter is at the instruction following the jal ({\em return integrity})
\end{itemize}

Non-skipping may seem strange -- surely the depth counter ought to decrease by one at a time
as part of the property definition, and the problem comes when the return is to a shallower
call's return pointer, violating return integrity? Yet some weaker enforcement mechanisms,
such as those in which returns are implemented as invocation of a Cheri object capability,
do allow skipping when adversarial code acts unreasonably by leaking its own return capability,
reasoning that this is acceptable as long as only the unreasonable function is skipped
(e.g. \citep{Skorstengaard+19b}).\apt{???} A variant of our model supports this weaker form of WBCF,
but with tag-based policies we can enforce the non-skipping version entirely.\apt{??}
\sna{Cite future work for details when the relevant section is finished.}

@@

\newcommand{\mainpassc}{magenta}
\newcommand{\sharec}{lgray}

\begin{figure}[b]
  \centering
  \begin{subfigure}{.45\textwidth}
\begin{verbatim}
int stash;

int main() {
0:             int x = 42, y = 0;
1[P(x)],2[C]:  f(&x);
3:             x = 0;
4[P(y)],5[C]:  f(&y);
6[R]:          return 0;
}

void f(int* a) {
10:  if(!stash)
11:    stash = a;
12:  *stash = 5;
13:  return;
}
\end{verbatim}
\end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l}
      {\tt 2} &
      \memoryaddrs{12em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
      ~$\cdots$ \\
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{2}{\unsealc}
      ~$\cdots$ \\
      {\tt 3} &
      \memoryaddrs{12em}
      \memory{3}{\unsealc}
      ~$\cdots$ \\
      {\tt 5} &
      \memoryaddrs{12em}
      \memory{1}{\unsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{2}{\unsealc}
      ~$\cdots$ \\
\end{tabular}
\end{subfigure}

@@

Until now we have assumed that function arguments and returns are passed solely in registers.
This is an oversimplication, which we will now correct. There are important stack safety concerns
when passing arguments on the stack, as seen in example \ref{fig:passing}. In this example,
{\tt main} passes a reference to {\tt x}, but then clears {\tt x} and instead passes one to
{\tt y}. Before each call, when the address of a variable is taken, we add a new annotation
\(P(x)\) or \(P(y)\) identifying the memory to be passed, and our stack model marks it in
magenta. \apt{Why are there two line numbers assoicated with the two calls?}
Then at the call, the parts of the caller's frame that are not passed are sealed, as usual. Passed
memory also tracks which depth owns it, and it may only be accessed by the immediate callee
(at its depth plus one).

\apt{This needs to be talked through in more detail.}
But because {\tt f} stashes the original pointer to {\tt x}, it will overwrite {\tt x} while
{\tt x} is sealed, an integrity violation. This fits with the intuition of passing by reference,
that the passed variable might be read or written during the call, not that it is permanently
open to be read or written by all subsequent calls. This also applies to passing large objects
by value in the stack: the caller allocates them, and the relevant instructions are annotated to
indicate passing.

We do not model cases where permanent sharing is desireable, because they are uninteresting
and hard to motivate: we would expect to treat such sharing as completely unprotected, and
perhaps attach a separate memory safety policy to add some protection to such variables.

>>>>

It may be unclear in figure \ref{fig:conf1} why we also vary the initial values of {\tt y}
and {\tt z}. In figure \ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
{\tt y} without even directly using its ability to read the stack pointer. In
this case, {\tt g} branches on that value, which is a different way of
violating lockstep confidentiality. Technically, this violation
occurs because the program counter is also a component of the state that must change
identically between the primary and variant traces. In a case where \(v_1 \not = 5\),
the program counters will differ, breaking the lockstep.

\sna{Not terribly happy with the circles yet; the idea is that instead of a meta-level
  label for the line numbers we could treat the PC as a part of the system we're showing.}

%\paragraph*{Control Flow and Awkward Examples}

%\begin{figure}
%  \centering
%  \begin{minipage}{.4\textwidth}
%\begin{verbatim}
%int stash = 0;

%int main() {
%0:     int x = 1;
%1[C]:  f();
%2:     x = -x;
%3[C]:  f();
%4[R]:  return x;
%}
%\end{verbatim}
%  \end{minipage}
%  \begin{minipage}{.4\textwidth}
%\begin{verbatim}
%void f() {
%10:     if(!stash) {
%        // store return pointer
%11:       stash = $SP-1;
%12:     } else {
%13:       $SP-1 = stash;
%14:       stash = 0;
%15[R]:    return;
%        }
%}
%\end{verbatim}
%  \end{minipage}
%\label{fig:WBCF1}
%  \begin{subfigure}[t]{.4\textwidth}
%\begin{verbatim}
%int counter = 0;
%int stash = 0;
%
%int main() {
%0:     int x = 1;
%1[C]:  f();
%2:     x = -x;
%3[C]:  f();
%4[R]:  return x;
%}
%
%void f() {
%10:     switch(counter++) {
%11:       case 0:
%12[R]:      return;
%13:       case 1:
%14:         stash = $SP-1;
%15[C]:      main();
%16:         return;
%17:       default:
%18:         $SP-1 = stash;
%19[R]:      return;
%        }
%}
%\end{verbatim}
%\label{fig:WBCF2}
%  \end{subfigure}
%  \caption{Control Flow Attacks}
%  \label{fig:WBCF}
%\end{figure}

%Example \ref{fig:WBCF} shows how control flow attacks can compromise integrity in principle
%even when the attacker obeys LSE. In this case, the attacker hides a return pointer in a
%global variable and retrieves it the second time it is called, returning to the first call
%point rather than the second. A return was performed, so {\tt x} is no longer sealed, but
%line {\tt 2} is executed twice, so that {\tt main} returns 1 instead of -1.

%To prevent this, we define Well-bracketed Control Flow. Throughout the above examples, we
%have implicitly tracked the depth of the current call in order to seal memory at the correct
%depth. We also have a clear sense of where the entry points to functions are -- the top line,
%in our examples, and another instruction annotation in practice. Well-bracketed Control Flow
%holds if:

%\begin{itemize}
%\item All calls jump to entry points ({\em entry integrity})
%\item The depth counter decreases by exactly one on a return ({\em non-skipping})
%\item From a call at depth \(d\), the next state to have depth \(d\) matches the stack pointer,
%  and the program counter is at the instruction following the jal ({\em return integrity})
%\end{itemize}

%Non-skipping may seem strange -- surely the depth counter ought to decrease by one at a time
%as part of the property definition, and the problem comes when the return is to a shallower
%call's return pointer, violating return integrity? Yet some weaker enforcement mechanisms,
%such as those in which returns are implemented as invocation of a Cheri object capability,
%do allow skipping when adversarial code acts unreasonably by leaking its own return capability,
%reasoning that this is acceptable as long as only the unreasonable function is skipped
%(e.g. \citep{Skorstengaard+19b}).\apt{???} A variant of our model supports this weaker form of WBCF,
%but with tag-based policies we can enforce the non-skipping version entirely.\apt{??}
%\sna{Cite future work for details when the relevant section is finished.}

@@

\newcommand{\mainpassc}{magenta}

\begin{figure}
  \centering
  \begin{subfigure}{.45\textwidth}
\begin{verbatim}
int stash = 0;

int main() {
0:         int x, y;
1[P(x)C]:  f(x);
2:         x = 0;
3[P(y)C]:  f(y);
4[R]:      return x;
}

void f(int &a) {
10:    if(!stash)
11:      stash = &a;
12:    *stash = 5;
13[R]:   return;
}
\end{verbatim}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l}
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 11} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}
      \memory{1}{\mainsealc}
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{5}
      \\
      {\tt 3} &
      \memoryaddrs{12em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 11} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{\bf 5}
\end{tabular}
\end{subfigure}

@@

Up til now we have ignored function arguments, since our simplest model places them in
registers. This is an oversimplication, which we will now correct, starting with the more
interesting case: call-by-reference. There are important stack safety concerns
when passing arguments by reference, as seen in example \ref{fig:passing}. In this example,
{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
{\tt x} in the second call.

At each call, we add a new annotation \(P(x)\) or \(P(y)\) identifying the memory to be
passed, and our stack model marks it in magenta. Then the caller's frame that are not
passed are sealed, as usual. Just as we marked sealed memory with the depth of its owner,
we do the same to passed memory. We keep track of the current depth of the active function
next to the stack pointer in the diagram. A passed address may only be accessed when the
current depth is one higher than the one it is sealed with.

So at the initial call, {\tt x} is passed, and {\tt f} is free to write to it. Then upon
return, {\tt main}'s frame is unsealed. At the second call it is {\tt y} that is passed,
and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
a future call with no arguments passed by reference would also violate integrity if it wrote
to either variable.

Passed addresses are also excluded from confidentiality: at the start of each call, we vary
only stack addresses that are not passed. This is important, as we use the passing label
for those arguments that are passed by value on the stack, as well. We might instead
treat them as unsealed, but then we would treat them as confidential like the rest of the stack.

We do not model permanent sharing, in which a variable's address is taken and it might be
passed far afield before use, because this situation is uninteresting and hard to motivate
as an instance of stack safety. We would expect to treat such sharing as completely
unprotected, and perhaps attach a separate memory safety policy to add some protection
to such variables.
